Surrogate model for aligned-spin binary neutron star waveforms using gaussian process regression

### 2017-07-18 paper/paper.tex
Tweak EOB formulas in paper

References~\cite{Hinderer:2016eia,Steinhoff:2016rfi} built upon the aligned-spin BBH EOB model of Ref.~\cite{Taracchini:2013rva} and proposed a way to include the effect of dynamical tides. Neutron stars that are part of a compact-object binary will deform in the tidal field generated by the companion --- be it a black hole or another neutron star. Because of binary motion, the forcing tidal field varies at the orbital frequency. Thus, in the late stages of the inspiral, the characteristic $f$-mode frequency of the neutron star can be dynamically approached, resulting in a resonant excitation of such mode. The net effect is an amplification of tidal effects as compared to the adiabatic limit, which assumes that the $f$-mode frequency is much larger than the frequency of the forcing tidal field. Dynamical tidal effects are implemented in the EOB model through a modification of the radial potential $\Delta_u$, which is the $tt$-component of the metric of the effective spacetime. Here we adopt the tidally-augmented expression of $\Delta_u$ that is discussed in Appendix~A of Ref.~\cite{Steinhoff:2016rfi}, namely
\begin{equation}
\Delta_u = \Delta_u^{\textrm{pm}} + \Delta_u^{\textrm{DT}}\,,
\end{equation}
where $\Delta_u^{\textrm{pm}}$ is the 4PN-accurate point-mass term of the underlying BBH EOB model of Ref.~\cite{Bohe:2016gbl} (see Eq.~(2.2) therein) and $\Delta_u^{\textrm{DT}}$ is the contribution due to dynamical tides. We consider the general case of a binary that may consist of two neutron stars; if either component of the binary is a black hole, then vanishing values of the tidal polarizabilities are employed in the formulas below. Let $A$ and $B$ be the body labels. Let $M^{A,B}$ be the mass of either body and $M=M^A+M^B$ be the total mass of the binary.  Explicitly, including quadrupolar and octupolar tidal effects, we have
\begin{widetext}
\begin{align}
 \Delta_u^{\textrm{DT}} =&  - 3\,\Lambda_{2,\textrm{dyn}}^{A}(u)\,u^6\frac{X_B}{X_A} \left[ 1 + \frac{5}{2} X_Au + \left(3 + \frac{1}{8}X_A + \frac{337}{28} X_A^2\right)u^2 \right]\nonumber\\
 &- 15\,\Lambda_{3,\textrm{dyn}}^{A}(u)\,u^8 \frac{X_B}{X_A}\left[1 + \left(-2 + \frac{15}{2}X_A\right)u + \left(\frac{8}{3} - \frac{311}{24}X_A + \frac{110}{3}X_A^2\right) u^2\right] + (A \leftrightarrow B)\,,
\end{align}
\end{widetext}
where $X_{A,B}=M^{A,B}/M$, $u=1/r$ is the inverse of the ($M$-rescaled) EOB radial coordinate, $\Lambda_{2,\textrm{dyn}}^{A,B}(u)$ are the dimensionless quadrupolar dynamical tidal polarizabilities, and $\Lambda_{3,\textrm{dyn}}^{A,B}(u)$ are the dimensionless octupolar dynamical tidal polarizabilities. Within the dynamical tides model, the tidal polarizabilities are not constant, but rather depend on the orbital separation and on the values of the $f$-mode frequencies. In particular, the dimensionless dynamical polarizabilities read
\begin{equation}
\Lambda_{\ell,\textrm{dyn}}^{A,B}(u)=\frac{2}{(2\ell-1)!!}k^{A,B}_{\ell}\hat{k}^{A,B}_{\ell\,\textrm{dyn}}(u;\omega_{0\ell}^{A,B})\left(\frac{R_{A,B}}{M}\right)^{2\ell+1}\,,
\end{equation}
where $k^{A,B}_{\ell}$ is the tidal Love number and $R_{A,B}$ is the neutron star radius. Here $\hat{k}^{A,B}_{\ell\,\textrm{dyn}}(u;\omega_{0\ell}^{A,B})$ is the separation-dependent, dimensionless enhancement factor given in Eq.~(11) of Ref.~\cite{Dietrich:2017feu}, which depends on the value of the $f$-mode angular frequency of the $2^\ell$-th multipole, $\omega_{0\ell}^{A,B}$.

>>>>

References~\cite{Hinderer:2016eia,Steinhoff:2016rfi} built upon the aligned-spin BBH EOB model of Ref.~\cite{Taracchini:2013rva} and proposed a way to include the effect of dynamical tides. Neutron stars that are part of a compact-object binary will deform in the tidal field generated by the companion --- be it a black hole or another neutron star. Because of binary motion, the forcing tidal field varies at the orbital frequency. Thus, in the late stages of the inspiral, the characteristic $f$-mode frequency of the neutron star can be dynamically approached, resulting in a resonant excitation of such mode. The net effect is an amplification of tidal effects as compared to the adiabatic limit, which assumes that the $f$-mode frequency is much larger than the frequency of the forcing tidal field. Dynamical tidal effects are implemented in the EOB model through a modification of the radial potential $\Delta_u$, which is the $tt$-component of the metric of the effective spacetime. Here we adopt the tidally-augmented expression of $\Delta_u$ that is discussed in Appendix~A of Ref.~\cite{Steinhoff:2016rfi}, namely
\begin{equation}
\Delta_u = \Delta_u^{\textrm{pm}} + \Delta_u^{\textrm{DT}}\,,
\end{equation}
where $\Delta_u^{\textrm{pm}}$ is the 4PN-accurate point-mass term of the underlying BBH EOB model of Ref.~\cite{Bohe:2016gbl} (see Eq.~(2.2) therein) and $\Delta_u^{\textrm{DT}}$ is the contribution due to dynamical tides. We consider the general case of a binary that may consist of two neutron stars; if either component of the binary is a black hole, then vanishing values of the tidal polarizabilities are employed in the formulas below. Let $A$ and $B$ be the body labels. Let $M_{A,B}$ be the mass of either body and $M=M_A+M_B$ be the total mass of the binary.  Explicitly, including quadrupolar and octupolar tidal effects, we have
\begin{widetext}
\begin{align}
 \Delta_u^{\textrm{DT}} =&  - 3\,\Lambda_{2,\textrm{dyn}}^{A}(u)X_A^4 X_B\,u^6  \left[ 1 + \frac{5}{2} X_Au + \left(3 + \frac{1}{8}X_A + \frac{337}{28} X_A^2\right)u^2 \right]\nonumber\\
 &- 15\,\Lambda_{3,\textrm{dyn}}^{A}(u)X_A^6X_B\,u^8 \left[1 + \left(-2 + \frac{15}{2}X_A\right)u + \left(\frac{8}{3} - \frac{311}{24}X_A + \frac{110}{3}X_A^2\right) u^2\right] + (A \leftrightarrow B)\,,
\end{align}
\end{widetext}
where $X_{A,B}=M_{A,B}/M$, $u=1/r$ is the inverse of the ($M$-rescaled) EOB radial coordinate, $\Lambda_{2,\textrm{dyn}}^{A,B}(u)$ are the dimensionless quadrupolar dynamical tidal polarizabilities, and $\Lambda_{3,\textrm{dyn}}^{A,B}(u)$ are the dimensionless octupolar dynamical tidal polarizabilities. Within the dynamical tides model, the tidal polarizabilities are not constant, but rather depend on the orbital separation and on the values of the $f$-mode frequencies. In particular, the dimensionless dynamical polarizabilities read
\begin{equation}
\Lambda_{\ell,\textrm{dyn}}^{A,B}(u)=\frac{2}{(2\ell-1)!!}\frac{k^{A,B}_{\ell}}{C_{A,B}^{2\ell+1}}\hat{k}_{\ell\,\textrm{dyn}}(u;\omega_{0\ell}^{A,B})\,,
\end{equation}
where $k^{A,B}_{\ell}$ is the tidal Love number and $C_{A,B}=R_{A,B}/M_{A,B}$ is the neutron star compactness, with $R_{A,B}$ the neutron star radius. Here $\hat{k}_{\ell\,\textrm{dyn}}(u;\omega_{0\ell}^{A,B})$ is the separation-dependent, dimensionless enhancement factor given in Eq.~(11) of Ref.~\cite{Dietrich:2017feu}, which depends on the value of the $f$-mode $2^\ell$-pole angular frequency, $\omega_{0\ell}^{A,B}$.
